Upper limits on CN from exocomets transiting $\beta$~Pictoris

### 2025-04-18 src/tex/ms.tex
tweaks
deps for analyze script

One of the closest young planetary systems is $\beta$~Pictoris, which has been studied intensively both with direct imaging and with spectroscopy.
%
\bp{} is a young \citep[$\sim 25$ Myr; see ][ for discussions]{Lee24}, nearby \citep[$19.63\pm 0.06$ pc, ][]{Lindegren21} and early type star \citep[A6V; ][]{Gray06} which hosts an edge-on debris disk and at least two exoplanets, making it one of the most intensely studied young planetary systems.
%
The star is a rapid rotator \citep[$124\pm 3$\,\kms{}; ][]{Koen03} seen close to equator on.
%
Amongst these, the Ca H and K lines show deep absorption down to 0.1 of the continuum level.
%
Deep within this rotationally broadened line there is a consistent and stable narrow absorption feature with a full width half maximum (FWHM) of a few km/s, attributed to the presence of the circumstellar material \citep{Hobbs85,VidalMadjar86}.

@@

\subsection{Removing the stellar spectrum}\label{sect:starcor}
The processing up to this stage removes most of the variations in the effective blaze function, but the overall stellar spectrum and, for HARPS, shape of the blaze profile, are still present and need to be removed.

@@

%\subsubsection{\ce{CN} Cross-section}
The effective cross section of the coma will depend on how much is projected against the disk of \bp{}.
%
The radius of the \ce{CN} coma will be given by
$$r\sim v_{\ce{CN}} \tau_{\ce{CN}}$$
The outflow velocity of the parent \ce{HCN} will depend on the nuclear surface temperature, which will increase as a comet moves towards its parent star.
%
\citet{Biver2002} found a nuclear gas outflow velocity $v\simeq1.0  r_h^{-0.4}$ \kms{} from sub-mm measurements of Comet Hale-Bopp, close to the canonical relationship of $r_h^{-0.5}$.
%
\citet{Harris2002} notes lower activity similar to Halley results in smaller outflow velocities.
%
The excess photodissociation velocity of \ce{CN} will not change.
%
Therefore gas outflow speeds are expected to be $\sim 1-2$ \kms{} at 1 AU, and the \ce{CN} coma radius will be
%
$$r\sim (1-2) \times 3000 \leq 6\times 10^{3}\:{\rm km}$$
%
This is much smaller than a stellar radius, so we can assume for an exoHalley that all of the coma is projected into the stellar disk as seen from Earth.

@@

Finally, comets have impulsive outbursts, where the amount of material ejected into the coma increases significantly over a few hours or less.
%
The general causes are unclear, but there is evidence for several physical mechanisms that cause them 
\citep{Kelley2021,Muller2024}.
``Normal'' outbursts generally result in a brightness increases of 1--5 magnitudes, with smaller outbursts being more common.
%
As the apparent brightness is directly proportional to the amount of material (gas+dust) in the coma, this implies short term increases in $N_{\ce{CN}}$ and $n$ of factors $\sim 2-100$.

@@

The authors thank the Lorentz Centre at Leiden University for organising the workshop ``Exocomets: Understanding the Composition of Planetary Building Blocks'' which led to the study in this paper. 
%
Part of this work was supported by the German \emph{Deut\-sche For\-schungs\-ge\-mein\-schaft, DFG\/} project number Ts~17/2--1.
%
AF acknowledges support from STFC award ST/T00021X/1. EdM acknowledges support from STFC award ST/X00094X/1.
%
We made use of the {\tt Python} programming language \citep{rossum1995} and the open-source {\tt Python} packages {\tt numpy} \citep{walt2011}, {\tt matplotlib} \citep{hunter2007}, {\tt astropy} \citep{astropy2013}.
%      
An online repository with materials used in this work is available at \url{https://github.com/mkenworthy/BetaPicCN}

@@

\section{Selected exocomets.}\label{sec:app_FEBs}
The dates for the nights selected to have the strongest exocomet features, as well as the depths and velocities of these features can be found in Tab.~\ref{tab:FEB_data}.
\begin{table}
\caption{Table with the nights with the strongest exocomet features. Note that the dates given are for the average of all the observations that were combined during each of the nights.}\label{tab:FEB_data}
\begin{tabular}{c c c}
   BJD-2,400,000  &    v$_{comet}$  &  Depth\\
 & [km/s] & \\
 \hline
\hline
53031.2 & 19.7 & 0.63\\
53032.2 & 18.9 & 0.74\\
53033.1 & 22.3 & 0.76\\
53036.1 & 15.5 & 0.67\\
53045.1 & 18.0 & 0.72\\
53046.1 & 18.0 & 0.91\\
53047.1 & 20.6 & 0.84\\
53048.0 & 20.6 & 0.83\\
53049.1 & 20.6 & 0.94\\
53050.1 & 18.9 & 0.90\\
53051.1 & 18.9 & 0.87\\
53052.1 & 18.9 & 0.92\\
53053.1 & 18.0 & 0.93\\
53054.1 & 17.2 & 0.83\\
53056.1 & 16.3 & 0.86\\
53057.1 & 15.5 & 0.93\\
53058.1 & 14.6 & 0.92\\
53059.1 & 14.6 & 0.87\\
53060.1 & 13.7 & 0.82\\
53061.1 & 13.7 & 0.72\\
53062.1 & 12.9 & 0.69\\
53063.1 & 12.0 & 0.67\\
53064.1 & 12.0 & 0.61\\
53581.4 & -8.5 & 0.65\\
54223.0 &  8.6 & 0.62\\
54542.1 & 29.2 & 0.91\\
54544.1 & 28.3 & 0.92\\
54545.0 & 23.2 & 0.79\\
54546.0 &  8.6 & 0.80\\
54799.2 &  8.6 & 0.75\\
54878.1 & 14.6 & 0.76\\
54913.0 & 12.0 & 0.81\\
56524.4 & 31.7 & 0.76\\
56534.3 & 12.9 & 0.76\\
56536.3 &  8.6 & 0.61\\
57849.0 & 10.3 & 0.73\\
\end{tabular}
\end{table}

>>>>

One of the closest young planetary systems is $\beta$~Pictoris, which has been studied intensively both with direct imaging and with spectroscopy.
%
\bp{} is a young \citep[$\sim 25$ Myr; see ][ for discussions]{Lee24}, nearby \citep[$19.63\pm 0.06$ pc, ][]{Lindegren21} and early type star \citep[A6V; ][]{Gray06} which hosts an edge-on debris disk and at least two exoplanets, making it one of the most intensely studied young planetary systems.
%
The star is a rapid rotator \citep[$124\pm 3$\,\kms{}; ][]{Koen03} seen close to equator on.
%
Amongst these, the \ce{Ca} H and K lines show deep absorption down to 0.1 of the continuum level.
%
Deep within this rotationally broadened line there is a consistent and stable narrow absorption feature with a full width half maximum (FWHM) of a few km/s, attributed to the presence of the circumstellar material \citep{Hobbs85,VidalMadjar86}.

@@

\subsection{Removing the stellar spectrum}\label{sect:starcor}

The processing up to this stage removes most of the variations in the effective blaze function, but the overall stellar spectrum and, for HARPS, shape of the blaze profile, are still present and need to be removed.

@@

The effective cross section of the coma will depend on how much is projected against the disk of \bp{}.
%
The radius of the \ce{CN} coma will be given by
$$r\sim v_{\ce{CN}} \tau_{\ce{CN}}$$
The outflow velocity of the parent \ce{HCN} will depend on the nuclear surface temperature, which will increase as a comet moves towards its parent star.
%
\citet{Biver2002} found a nuclear gas outflow velocity $v\simeq1.0  r_h^{-0.4}$ \kms{} from sub-mm measurements of Comet Hale-Bopp, close to the canonical relationship of $r_h^{-0.5}$.
%
\citet{Harris2002} notes lower activity similar to Halley results in smaller outflow velocities.
%
The excess photodissociation velocity of \ce{CN} will not change.
%
Therefore gas outflow speeds are expected to be $\sim 1-2$ \kms{} at 1 AU, and the \ce{CN} coma radius will be
%
$$r\sim (1-2) \times 3000 \leq 6\times 10^{3}\:{\rm km}$$
%
This is much smaller than a stellar radius, so we can assume for an exoHalley that all of the coma is projected into the stellar disk as seen from Earth.

@@

Finally, comets have impulsive outbursts, where the amount of material ejected into the coma increases significantly over a few hours or less.
%
The general causes are unclear, but there is evidence for several physical mechanisms that cause them \citep{Kelley2021,Muller2024}.
%
``Normal'' outbursts generally result in a brightness increases of 1--5 magnitudes, with smaller outbursts being more common.
%
As the apparent brightness is directly proportional to the amount of material (gas+dust) in the coma, this implies short term increases in $N_{\ce{CN}}$ and $n$ of factors $\sim 2-100$.

@@

The authors thank the Lorentz Centre at Leiden University for organising the workshop ``Exocomets: Understanding the Composition of Planetary Building Blocks'' which led to the study in this paper. 
%
Part of this work was supported by the German \emph{Deut\-sche For\-schungs\-ge\-mein\-schaft, DFG\/} project number Ts~17/2--1.
%
AF acknowledges support from STFC award ST/T00021X/1.
%
EdM acknowledges support from STFC award ST/X00094X/1.
%
We made use of the {\tt Python} programming language \citep{rossum1995} and the open-source {\tt Python} packages {\tt numpy} \citep{walt2011}, {\tt matplotlib} \citep{hunter2007}, {\tt astropy} \citep{astropy2013}.
%      
An online repository with materials used in this work is available at \url{https://github.com/mkenworthy/BetaPicCN} using the {\tt showyourwork!} package \citep{Luger2021}.

@@

\section{Selected exocomets}\label{sec:app_FEBs}
The dates for the nights selected to have the strongest exocomet features, as well as the depths and velocities of these features can be found in Table~\ref{tab:FEB_data}.
\begin{table}[h]
\caption{Nights with the strongest exocomet features.}
\label{tab:FEB_data}
\begin{tabular}{c c c}
   BJD-2,400,000\tablefootmark{a}  &    v$_{comet}$  &  Depth\\
 & [km/s] & \\
 \hline
\hline
53031.2 & 19.7 & 0.63\\
53032.2 & 18.9 & 0.74\\
53033.1 & 22.3 & 0.76\\
53036.1 & 15.5 & 0.67\\
53045.1 & 18.0 & 0.72\\
53046.1 & 18.0 & 0.91\\
53047.1 & 20.6 & 0.84\\
53048.0 & 20.6 & 0.83\\
53049.1 & 20.6 & 0.94\\
53050.1 & 18.9 & 0.90\\
53051.1 & 18.9 & 0.87\\
53052.1 & 18.9 & 0.92\\
53053.1 & 18.0 & 0.93\\
53054.1 & 17.2 & 0.83\\
53056.1 & 16.3 & 0.86\\
53057.1 & 15.5 & 0.93\\
53058.1 & 14.6 & 0.92\\
53059.1 & 14.6 & 0.87\\
53060.1 & 13.7 & 0.82\\
53061.1 & 13.7 & 0.72\\
53062.1 & 12.9 & 0.69\\
53063.1 & 12.0 & 0.67\\
53064.1 & 12.0 & 0.61\\
53581.4 & -8.5 & 0.65\\
54223.0 &  8.6 & 0.62\\
54542.1 & 29.2 & 0.91\\
54544.1 & 28.3 & 0.92\\
54545.0 & 23.2 & 0.79\\
54546.0 &  8.6 & 0.80\\
54799.2 &  8.6 & 0.75\\
54878.1 & 14.6 & 0.76\\
54913.0 & 12.0 & 0.81\\
56524.4 & 31.7 & 0.76\\
56534.3 & 12.9 & 0.76\\
56536.3 &  8.6 & 0.61\\
57849.0 & 10.3 & 0.73\\
\hline
\end{tabular}
\tablefoot{\tablefoottext{a}{Dates given are for the average of all the observations that were combined during each of the nights.}}
\end{table}
